|

### 2023-02-06 paper.Rmd
Mon Feb  6 17:00:09 CET 2023: update-remote.sh

---
title: |
  <center>Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming</center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Master Thesis - M.Sc. Business Education</center>
  <center>FB 03: Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    toc_depth: 3
    includes:
      in_header: [header_files/settings.tex]
indent: false
fontsize: 12pt
bibliography: ./ref/ref.bib
link-citations: true
fig.width: 4in
fig.height: 3in
csl: csl/association-for-computing-machinery.csl
# csl: csl/harvard-cite-them-right.csl
# csl: csl/springer-science-reviews.csl
urlcolor: blue
linkcolor: blue
toccolor: blue
github-repo: https://github.com/roman91DE/master_thesis
include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
---

@@

The framework that was used to conduct all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation.

@@

```{r airfoil-runtime-noSampling-total, echo=FALSE, fig.cap="Median Runtime excluding Time for Sampling - Airfoil", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling.png")
```
While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`). After profiling a single run of DAE-GP and pre-trained DAE-GP my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training. Figure  \@ref(fig:airfoil-profiling) shows the cumulative time spend for the 20 most expensive function calls. Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to `DAE_LSTM.py:208(sample)`) than during model training (calls to `training.py:1303(fit)`). Secondly, it shows that internal error handling of the baseline neural network library `keras` (calls to `traceback_utils.py:138(error_handler)` and `traceback_utils.py:59(error_handler)`) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
    }
  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
  }
}
# shorten column names
col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
    }
  else if (df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
  }
}
# shorten column names
col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
    }
  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
  }
}
# shorten column names
col_names <- c("Problem", "Hid.Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

<!--
  EIDESSTAATLICHE ERKLAERUNG
# II - Statutory Declaration {.unnumbered #II}

![Statutory Declaration](./private/erklaerung.pdf)\
-->

>>>>

---
title: |
  <center>Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming</center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Master Thesis - M.Sc. Business Education</center>
  <center>FB 03: Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    toc_depth: 3
    includes:
      in_header: [header_files/settings.tex]
indent: false
fontsize: 11pt
bibliography: ./ref/ref.bib
link-citations: true
fig.width: 4in
fig.height: 3in
csl: csl/association-for-computing-machinery.csl
# csl: csl/harvard-cite-them-right.csl
# csl: csl/springer-science-reviews.csl
urlcolor: blue
linkcolor: blue
toccolor: blue
github-repo: https://github.com/roman91DE/master_thesis
include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
---

@@

The framework that was used to run all experiments of this thesis was provided by the supervisor of this thesis David Wittenberg and uses the python programming language [@python3] in conjunction with the baseline libraries `Keras` [@chollet2015keras] for deep neural networks and `deap` [@DEAP_JMLR2012] for evolutionary computation. For data visualizations I used the python package `matplotlib` [@matplotlib], statistical analysis and evaluations were conducted using primarily the datscience libraries `numpy` [@numpy], `pandas` [@pandas], and

@@

```{r airfoil-runtime-noSampling-total, echo=FALSE, fig.cap="Median Runtime excluding Time for Sampling - Airfoil", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_150hn_fullRun_30gens_TimeBenchmark/median_runtime_withoutSampling.png")
```
While investigating other reasons for the differences in run-time, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models (total number of trainable parameters DAE-GP=`r format(REG_PARAMS, scientific=F)`, Pre-trained DAE-GP=`r format(PT_PARAMS, scientific=F)`). After profiling a single run of DAE-GP and pre-trained DAE-GP using the python `cProfiler` module, my best explanation for the remaining differences in run-time is an inefficiency in my implementation of pre-training. Figure  \@ref(fig:airfoil-profiling) shows the cumulative time spend for the 20 most expensive function calls. Even though it is difficult to identify the exact reasons for the performance hit for pre-trained DAE-GP, it shows two interesting findings: Firstly, both algorithms spend relatively more time during model sampling (calls to `DAE_LSTM.py:208(sample)`) than during model training (calls to `training.py:1303(fit)`). Secondly, it shows that internal error handling of the baseline neural network library `keras` (calls to `traceback_utils.py:138(error_handler)` and `traceback_utils.py:59(error_handler)`) takes up a comparatively larger amount of run time for pre-trained DAE-GP than for regular DAE-GP.

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],4)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],4))
    }
  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],4)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],4))
  }
}
# shorten column names
col_names <- c("Problem", "Hidden Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
    }
  else if (df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
  }
}
# shorten column names
col_names <- c("Problem", "Hidden Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

for(i in 1:nrow(df)){
  if(df[i, "DAE.GP"] < df[i, "Pre.Trained_DAE.GP"]){
    df_bold[i, "DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "DAE.GP"],2)))
    df_bold[i, "Pre.Trained_DAE.GP"] <- as.character(round(df[i, "Pre.Trained_DAE.GP"],2))
    }
  else if (df[i, "DAE.GP"] > df[i, "Pre.Trained_DAE.GP"]) {
    df_bold[i, "Pre.Trained_DAE.GP"] <- sub("^(.*)$", "**\\1**", as.character(round(df[i, "Pre.Trained_DAE.GP"],2)))
  df_bold[i, "DAE.GP"] <- as.character(round(df[i, "DAE.GP"],2))
  }
}
# shorten column names
col_names <- c("Problem", "Hidden Layers", "DAE-GP", "Pre-Trained", "P-Value", "Cliffs-Delta")

@@

# II - Statutory Declaration {.unnumbered #IV}

...hier einfügen...
